Emergent dual scaling of riverine biodiversity

### 2021-03-24 coverletter.Rmd
Update coverletter.Rmd

In this research, we used theory and analysis of extensive fish data to show that the scale-invariant complexity of branching networks is an alternative dimension of ecosystem structure to dictate riverine biodiversity patterns. Theory and empirical analysis agreed that, **regardless of ecological traits of constituent species, riverine biodiversity exhibited a dual scaling law along the two orthogonal dimensions of ecosystem size and scale-invariant complexity.** An elegant consistency of theoretical and empirical results provided strong evidence that the dual scaling of biodiversity is a pervasive yet overlooked feature of complex ecosystems.

Our study distinguishes itself from previous studies in three important ways. (1) Our study is the first to show the *dual* scaling law of biodiversity. Notably, this pattern has emerged irrespective of ecological contexts or geographic regions, emphasizing the generality of our finding. (2) Our empirical analysis used *watershed ecosystems* as individual replicates of statistical analysis; our data set included 180 watersheds with 6605 sites across Japan and the US and this extensive data enabled us to perform the ecosystem-level analysis. Therefore, unlike most previous studies focused on local biodiversity patterns *within a single watershed*, this study provides crucial insights into how ecological communities are structured *across watersheds*. (3) The present study provides the foundation for complexity-based ecosystem management. While there have been substantial debates as to whether area-based ecosystem management is sufficient to conserve biodiversity at broad spatial scales (e.g., Akasaka et al, Conservation Letters, 10: 539-546), there is, to our knowledge, no conservation scheme that explicitly accounts for ecosystem complexity in the design. Hence, our research provides a novel perspective to spatial planning of protected areas or ecological restoration.

In addition to the significance described above, the connection between network complexity and biodiversity has a broad implication for a truly interdisciplinary audience, including scientists working on fluvial geomorphology, network theory, and biodiversity conservation. Our study is also timely given the unprecedented rate of global biodiversity loss in freshwater ecosystems. Therefore, *Proceedings of the National Academy Sciences* is an ideal outlet for this manuscript.

>>>>

In this research, we used theory and analysis of extensive fish data to show that the scale-invariant complexity of branching networks is an alternative dimension of ecosystem structure to dictate riverine biodiversity patterns. Theory and empirical analysis agreed that, **regardless of ecological traits of constituent species, riverine biodiversity exhibited a dual scaling law along the two orthogonal dimensions of ecosystem size and scale-invariant complexity.** The elegant consistency of theoretical and empirical results provided strong evidence that the dual scaling of biodiversity is a pervasive yet overlooked feature of complex ecosystems.

Our study distinguishes itself from previous studies in three important ways. (1) Our study is the first to show the *dual* scaling law of biodiversity. Notably, this pattern has emerged irrespective of ecological contexts or geographic regions, emphasizing the generality of our findings. (2) Our empirical analysis used *watershed ecosystems* as individual replicates of statistical analysis; our data set included 180 watersheds with 6605 sites across Japan and the US and this extensive data enabled us to perform the ecosystem-level analysis. Therefore, unlike most previous studies focused on local biodiversity patterns *within a single watershed*, this study provides crucial insights into how ecological communities are structured *across watersheds*. (3) The present study provides the foundation for complexity-based ecosystem management. While area-based ecosystem management spearheads global conservation efforts (e.g., protected areas), we are unaware of any conservation scheme that explicitly accounts for ecosystem complexity in the design. Hence, our research provides a novel perspective to spatial planning of protected areas or ecological restoration.

In addition to the significance described above, the connection between network complexity and biodiversity has a broad implication for a truly interdisciplinary audience, including scientists working on fluvial geomorphology, network theory, and biodiversity. Our study is also timely given the unprecedented rate of global biodiversity loss in freshwater ecosystems. Therefore, *Proceedings of the National Academy Sciences* is an ideal outlet for this manuscript.
